---
title: "Data_Cleaning"
output: html_document
---

# September 2020 class, Week 1, Day 2
Data cleaning exercise
Filename in R on workstation -- > 20210914_NCEAS_DSP_Wk1D2.R

## Clean workspace
```{r clear_workspace}
rm(list=ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SCM Notes:
Message = FALSE in chunk option will get the message to not display.
However, I learned something about the messages
Type out "package_name::function_name()" to get it to not default the first one it encounters (if the one you want is the 2nd one)
Code chunk shortcut: ctrl / alt / i
Pipe shortcut: ctrl / shift / m

# Install Packages
```{r library_packages, message = FALSE}
library(dplyr)
library(tidyr)
```

# Data sources
Load the cvs file from the website, right-click the download button to get the url
Data from [KNB](https://knb.ecoinformatics.org/view/df35b.304.2)
```{r, get_data}
data_catch <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

# Clean and reshape data
[1] Remove unnecessary columns
[2] Check column types
[3] Reshape data
[4] group and summarize data



[1] 
Remove unnecessary columns
Takes data_catch dataframe and then a selection of that makes the data_catch_clean dataframe.
Can select what you want (below)
But can also select what you want to drop
select(-All, -notesRegCode)
```{r, drop_unnecessary}
data_catch_clean <- data_catch %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

glimpse(data_catch_clean)
```
[2]
Check column types
Seems like Chinook is character but it needs to be integer

```{r, catch_clean}
catch_clean <- data_catch_clean %>% 
  mutate(Chinook = as.integer(Chinook))
```

Explore why there is an error
There is one, line 401, which is "I" and not "1"
```{r}
xx <- which(is.na(catch_clean$Chinook))
xx

xy <- which(is.na(data_catch_clean$Chinook))
  xy
```
Remove the error, replace the "I" with "1"
Replace I with the character 1 otherwise leave it as is
Then pipe that and change the Chinook value to integer
```{r}
catch_clean <- data_catch %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.integer(Chinook))

data_catch_clean <- catch_clean %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

glimpse(data_catch_clean)
```

[3] 
reshape data

Pivot longer, have wide table pivoting it to long
first argument, columns over which you want to pivot over (not region or year)
second argument = names to (column create out of column names) and values to (the actual catch data)

```{r}
catch_long <- data_catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to = "catch")

head(catch_long)
```

Pivot wider is opposite of longer
Just run through as an example
```{r}
  catch_wide <- catch_long %>%
  pivot_wider(names_from = Region, values_from = catch)

head(catch_wide)
```

Rename catch_long override it with another catch_long
This is "less brittle" than doing the rename in base R because it may break silently 
Then you may want to run mutate to get catch column that is actually number of fish
```{r}
catch_long <- catch_long %>%
  rename(catch_thousands = catch) %>%
  mutate(catch = catch_thousands * 1000)

head(catch_long)
```

[4]
group and summarize data

```{r}
mean_region <- catch_long %>%
  group_by(Region) %>%
  summarise(catch_mean = mean(catch_thousands))
```
```{r}
sd_region <- catch_long %>%
  group_by(Region) %>%
  summarise(std_region = sd(catch_thousands))
```




# Join to Regions Definitions




