---
title: "Functions_07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, clear_workspace}

rm(list=ls())

```


# Day 3 of Week 1, see the .R file on workstation -- > 

## Set up some test data
Provide a vector, calculate the Celcius value for each one of the F values and do it manually for all 4

```{r, load_data}

airtemps <- c(212, 78.3, 30, 32)

```

```{r, FtoC_poorly}

celcius1 <- (airtemps[1] - 32) * 5/9
celcius1
celcius2 <- (airtemps[2] - 32) * 5/9
celcius2
celcius3 <- (airtemps[3] - 32) * 5/9
celcius3
celcius4 <- (airtemps[4] - 32) * 5/9
celcius4

```


Provide a vector and do the same thing using a function, want it to take F value as input, create a C value as output (return value)
```{r, FtoC}

FtoC <- function(fahr) {
  celcius <- (fahr - 32) * 5/9
  return(celcius)
}
  
celcius5 <- FtoC(airtemps[1])
celcius5
celcius1 == celcius5

```
Show how to use functions to calculate vectorized values
```{r, FtoC_OnArray}

celcius6 <- FtoC(airtemps)
celcius6

```

Create my own function, C to F
```{r, CtoF}

CtoF <- function(celcius) {
  fahr <- (celcius*9/5) + 32
  return(fahr)
}

fahr <- CtoF(celcius6)
fahr
fahr == airtemps
```
Take F values as input but instead of just doing 1 thing do 2 things
[1] calculate C from F (same as before)
[2] calculate a kelvin value
[3] return multiple different values calculated within the function in a single array (telling sequence to call a lot of values -100 to 100)
```{r, FtoCtoK}

FtoCtoK <- function(fahr) {
  celcius <- (fahr - 32) * 5/9
  kelvin <- celcius + 273.15
  return(list(fahr = fahr, celcius = celcius, kelvin = kelvin))
}

temps_df <- data.frame(FtoCtoK(seq(-100, 100, 10)))
temps_df

```

Use function to perform parts of data analysis that create graphs and plots
Figure out styling and formatting you want for your plots so you can do it once and then apply it
Build scatter plot of F vs C data from above
If someone does not provide a base_size parameter then it will default to 9 as we request it
[1] turns off axis ticks
[2] makes the size of text to base-size you set
```{r, themes_plot}
themes_plot <- function(base_size = 9){
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    text = ggplot2::element_text(family = 'Helvetica', color = 'gray30', size = base_size)
  )
  
}

library(ggplot2)

ggplot(temps_df, mapping = aes(x = fahr, y = celcius, color = kelvin)) + 
  geom_point() + 
  themes_plot(10)


```


Scatter plot 
pass in a dataframe
2 new parameters the point size and the font-size passing in the dataframe temps_df
can use multiple variables or multiple arguments to customize 
```{r, scatterplot}
scatterplot <- function(df, point_size = 2, font_size=9) {
  ggplot(df, mapping=aes(x=fahr, y=celcius, color=kelvin)) +
    geom_point(size=point_size) +
    themes_plot(font_size)
}

scatterplot(temps_df, point_size = 3, font_size = 16)

```






```{r}


```







